Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-31 EconGrowth.Rnw
Now generating a comparison between predicted and actual GDP for all countries in a lattice graph.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countries <- c(US="US", UK="UK", JP="JP", CN="CN", ZA = "ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. "1" means first alphabetically.
# So, this list sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9)
# Full page lattice plot sizes and other parameters
layoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5 #Inches
figHeight <- 7.5 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 0.75 #75% of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
# List of energy types
energyTypes <- c(Q = "Q", X = "X", U = "U")
@

@@

<<Country Factors Graph, echo=FALSE>>=
createCountryFactorsGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "l"
  lineTypes <- c(1, 5, 2, 4, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 1, 1, 1, 1) #line widths
  colors <- c("black", "black", "black", "black", "black") #line and point colors
  lineSpec <- list(lty=lineTypes, lwd=lineWidths, col=colors)
  graph <- xyplot(iCapStk+iLabor+iQ+iX+iU ~ Year, data=dataTable,
                  type=graphType,
                  par.settings = list(superpose.line = lineSpec),
                  key=list(text=list(c("k", "l", "q", "x", "u")),
                           type=graphType,
                           lines=lineSpec,
                           columns=1, x=0.0, y=0.98),
                  scales=list(cex=1.0, #controls text size on scales
                              tck=scaleTickSize), #controls tick mark length
                  ylab="Indexed (1980=1)")
  return(graph)
}
@
<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, 
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryName){
  model <- cobbDouglasModel(countryName)
  coefs <- coef(usModel)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  return(data.frame(pred))
}

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  preds <- lapply(countries, cobbDouglasPredictions)
  out <- rbind(preds[["US"]], preds[["UK"]], preds[["JP"]], preds[["CN"]], preds[["ZA"]], preds[["SA"]], preds[["IR"]], preds[["TZ"]], preds[["ZM"]])
  colnames(out) <- c("predGDP")
  print(out)
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTable <- function(){
  dataCD <- rbind(cobbDouglasCountryRow("US"),
                cobbDouglasCountryRow("UK"),
                cobbDouglasCountryRow("JP"),
                cobbDouglasCountryRow("CN"),
                cobbDouglasCountryRow("ZA"),
                cobbDouglasCountryRow("SA"),
                cobbDouglasCountryRow("IR"),
                cobbDouglasCountryRow("TZ"),
                cobbDouglasCountryRow("ZM"))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countries
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas parameters for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc")
  return(tableCD)
}
@

@@

<<US_CD_Graph>>=
usModel <- cobbDouglasModel("US")
# print(coef(usModel))
usPred <- predict(usModel) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
#print(usPred)
data.frame(usPred)

# allData <- loadData("All")
# predictions <- cobbDouglasPredictionsColumn()
# allData <- cbind(allData, predictions)
# 
# print(allData)
@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas model results.", echo=FALSE>>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "y_pred")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

<<CDq analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countries, cobbDouglasEnergyTable, energyType="Q")
@

@@

<<CDx analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDxTables <- lapply(countries, cobbDouglasEnergyTable, energyType="X")
@

@@

<<CDu analysis, eval=FALSE>>=
CDuTables <- lapply(countries, cobbDouglasEnergyTable, energyType="U")
@

@@

#CESqTables <- lapply(countries, cesTable, energyType="Q")
@

>>>>

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.

# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. "1" means first alphabetically.
# So, this list sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9)
# Full page lattice plot sizes and other parameters
layoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5 #Inches
figHeight <- 7.5 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 0.75 #75% of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
# List of energy types
energyTypes <- c(Q="Q", X="X", U="U")
@

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"),
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryName){
  model <- cobbDouglasModel(countryName)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # There are two countries which do not have labor data for 2011. Need to add a "NA" for 2011 for those countries.
  if (countryName == "CN" || countryName == "ZA"){
    dfNA <- data.frame(NA)
    colnames(dfNA) <- "pred" #Need to give it the same name so that we can apply rbind
    df <- rbind(df, dfNA)
  }
  return(df)
}

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  preds <- lapply(countryAbbrevs, cobbDouglasPredictions)
  out <- rbind(preds[["US"]], preds[["UK"]], preds[["JP"]], preds[["CN"]], preds[["ZA"]], preds[["SA"]], preds[["IR"]], preds[["TZ"]], preds[["ZM"]])
  colnames(out) <- c("predGDP")
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTable <- function(){
  dataCD <- rbind(cobbDouglasCountryRow("US"),
                cobbDouglasCountryRow("UK"),
                cobbDouglasCountryRow("JP"),
                cobbDouglasCountryRow("CN"),
                cobbDouglasCountryRow("ZA"),
                cobbDouglasCountryRow("SA"),
                cobbDouglasCountryRow("IR"),
                cobbDouglasCountryRow("TZ"),
                cobbDouglasCountryRow("ZM"))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas parameters for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc")
  return(tableCD)
}
@

@@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas (without energy) results.", echo=FALSE>>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "Predicted")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,8), #y axis limits
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

<<CDq analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countryAbbrevs, cobbDouglasEnergyTable, energyType="Q")
@

@@

<<CDx analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDxTables <- lapply(countryAbbrevs, cobbDouglasEnergyTable, energyType="X")
@

@@

<<CDu analysis, eval=FALSE>>=
CDuTables <- lapply(countryAbbrevs, cobbDouglasEnergyTable, energyType="U")
@

@@

#CESqTables <- lapply(countryAbbrevs, cesTable, energyType="Q")
@
